---
title: "Assignment 1"
author: "Zoe Iacovino"
date: "9/6/2020"
output: html_document
---

## Loading 2017 ACS1
I decided to use 2017 in case I come back to a 2017 CPS Food Security set. <br>
```{r loading survey variables, results = 'hide', message=FALSE}
library(tidycensus)
library(tidyverse)
person_vars_2017 <- pums_variables %>% 
  distinct(year,survey,var_code, var_label,data_type,level) %>% 
  filter(year == 2017, survey == "acs1", level == "person")
hh_vars_2017 <- pums_variables %>%
  distinct(year,survey,var_code, var_label,data_type,level) %>%
  filter (year == 2017, survey == "acs1", level == "housing")

```

```{r tibbles}
person_vars_2017
hh_vars_2017
```

## Choosing Variables
I selected seven variables between person-level and household-level analysis to examine issues of food insecurity, sex, race, family composition, and recently pregnant individuals.
I selected seven variables between person-level and household-level analysis to examine issues of food insecurity, sex, race, family composition, and recently pregnant individuals. The resulting tibble has 51,564 observations of the 7 variables. 
<br>
<b> Person Data </b><br>
<ul>
<li>Age [AGEP] (Continuous)
<li>Gave birth in the last 12 months [FER] (Categorical)
<li>Sex [SEX] (Categorical)
<li>Race [RAC1P] (Categorical)
</ul>
<b> Household Data </b> <br>
<ul>
<li>Related Children in the Household [NRC] (Continuous)
<li>Receiving SNAP [FS] (Categorical)
<li>Family Income [FINCP] (Categorical)
</ul> 
<b> Compiling Data </b>
<br>
```{r Selecting variables, results = 'hide', message=FALSE}

ma_fam_assist<- get_pums(variables = c("NRC",
                                "FS",
                                "FINCP",
                                "SEX",
                                "FER",
                                "AGEP",
                                "RAC1P"
                                   ),
                      state = "MA",
                      year = 2017,
                      survey = "acs1",
                      recode = "TRUE") %>%
              filter(FINCP > 0) %>%
              select(NRC, FS_label, FINCP, SEX_label, FER_label,AGEP, RAC1P_label)

```


```{r}
ma_fam_assist
write.csv(ma_fam_assist, "2017 ACS Family Assistance.csv")
```


## Some Questions
 <ul>
 <li> How many pregnant individuals receiving supplemental assistance gave birth to a new baby?
 <li> How many new parents were income-eligible for SNAP but not receiving?
 <li> How many families would have been affected by the assistance cap in effect in MA in 2017? 
 <li> Is there any relationship between number of children in the household and family income?
 <li> Does race have a measurable affect on access to SNAP for eligible families? For recent new parents?
 <li> Does family income tend to increase with age? If so are there factors that disrupt that trend?
 </ul>
