---
title: "Assignment 1"
author: "Zoe Iacovino"
date: "9/6/2020"
output: html_document
---

## Loading 2017 ACS1
I decided to use 2017 in case I come back to a 2017 CPS Food Security set. <br>
```{r loading survey variables, results = 'hide', message=FALSE}
library(tidycensus)
library(tidyverse)
person_vars_2017 <- pums_variables %>% 
  distinct(year,survey,var_code, var_label,data_type,level) %>% 
  filter(year == 2017, survey == "acs1", level == "person")
hh_vars_2017 <- pums_variables %>%
  distinct(year,survey,var_code, var_label,data_type,level) %>%
  filter (year == 2017, survey == "acs1", level == "housing")

```


```{r tibbles}
person_vars_2017
hh_vars_2017
```

## Choosing Variables
I selected seven variables between person-level and household-level analysis to examine issues of food insecurity, sex, race, family composition, and recently pregnant individuals. The resulting tibble has 51,564 observations of the 7 variables. 
<br>
<b>Person Data</b>

<li>Age [AGEP] (Continuous)
<li>Gave birth in the last 12 months [FER] (Categorical)
<li><strike>Sex [SEX] (Categorical) </strike>
<li><strike> Race [RAC1P] (Categorical) </strike>
</ul>
<br>
<b> Household Data </b> <br>
<ul>
<li>Related Children in the Household [NRC] (Continuous)
<li>Receiving SNAP [FS] (Categorical)
<li>Household Income [HINCP] (Continuous)
<li> Number of People in Household [NP] *New*
</ul> 
<b> Compiling Data </b>
<br>

```{r Selecting variables, results = 'hide', message=FALSE}
#Update 9/21: Converting all variables to household level. Filtering out households with no related children,  
#First, I am pulling out my person level variables separately, making FER_label a binary,
person_data<- get_pums(variables = c(
                                "FER",
                                "AGEP"
                                   ),
                      state = "MA",
                      year = 2017,
                      survey = "acs1"
                      )%>%
              mutate(FER = ifelse(FER == "b", 2, FER))%>%
              mutate(FER = ifelse(FER == 2, 0, FER))%>%
              group_by(SERIALNO)%>%
              summarize(AGE = mean(AGEP),
                        FERH = max(FER)
                        )
              
person_data

#what does the 2 in FER mean? 
ugh <- get_pums(variables = "FER",
                      state = "MA",
                      year = 2017,
                      survey = "acs1",
                      recode = TRUE
                      )%>%
      select(FER, FER_label)
#Ah I see, instead of no = 0, they made no = 2. Rude. I will mutate all the b's to be 2, then mutate all the 2s to be 0s, so i can sum the household. Then, 1s will be households with a new baby and 0s will households without a new baby. 

#Turns out actually, R is smart enough to stop me from trying math on a categorical variable. I can however, assign the highest value in the household with the max() function.


 # pulling out select argument to see if I can get serial no select(SEX_label, FER_label, AGEP, RAC1P_label)%>% (It worked)

#household level data: 
#so after all that I think I can do these all together? If I have to group the Household together anyways? (It worked)
ma_fam_assist_raw<- get_pums(variables = c("NRC",
                                "FS",
                                "HINCP",
                                "NP",
                                "FER",
                                "AGEP"
                                   ),
                      state = "MA",
                      year = 2017,
                      survey = "acs1"
                      )%>%
                filter(NRC > 0, 
                       FS != "b",
                         ) %>%
                mutate(FER = ifelse(FER == "b", 2, FER))%>%
                mutate(FER = ifelse(FER == 2, 0, FER), 
                       HINCP = ifelse(HINCP < 0, 0, HINCP),
                       FS = ifelse(FS == 2, 0, FS))%>%
                group_by(SERIALNO)%>%
                summarize(
                      KIDS = mean(NRC),
                      SNAP = max(FS),
                      FINC = max (HINCP),
                      NPL = mean(NP),
                      AGE = mean(AGEP),
                      FERH = max(FER)) %>%
 select(KIDS, SNAP, FINC, NPL, AGE, FERH)

```
Hooray! Now I have a data frame fully on household level analysis. In this next section, I am recoding the binary SNAP and FERH variables to answery "Is the household recieving SNAP?" and "Did someone in the household give birth in the last 12 months?"


```{r Recoding my variables }
ma_fam_assist <- ma_fam_assist_raw %>%
  mutate(SNAP = ifelse(SNAP==0, "No", "Yes"),
         FERH = ifelse(FERH==0, "No", "Yes"))%>%
    select(KIDS, SNAP, FINC, NPL, AGE, FERH)

ma_fam_assist
```





```{r}
ma_fam_assist
write.csv(ma_fam_assist, "2017 ACS Family Assistance Household.csv")
```


## Some Questions
 <ul>
 <li> How many pregnant individuals receiving supplemental assistance gave birth to a new baby?
 <li> How many new parents were income-eligible for SNAP but not receiving?
 <li> How many families would have been affected by the assistance cap in effect in MA in 2017? 
 <li> Is there any relationship between number of children in the household and family income?
 <li> Does race have a measurable affect on access to SNAP for eligible families? For recent new parents?
 <li> Does family income tend to increase with age? If so are there factors that disrupt that trend?
 </ul>



```{r}
ACS_Food_Insecurity <- get_pums(variables = "FS",
                      state = "MA",
                      year = 2017,
                      survey = "acs1"
                      )%>%
      select(FS, SERIALNO)

#write.csv(ACS_Food_Insecurity, "ACS1_2017_Food_Insecurity")
```

